---
title: "Central Limit Theorem"
author: "Michael Houck"
date: "August 23, 2015"
output: pdf_document
---

## Summary

The Central Limit Theorem states that, under certain cirsumstances, taking a sufficient number of sample means from a random variable is itself a random variable which will be normally distributed around the original population mean. This analysis will show this theory in pracice via simulation using the exponential distribution.

## Experiment

### Set-up

First, we will set up the experiment variables: the random number seed, the number of draws, the number of simulation iterations and lambda.

```{r experiment_setup}

set.seed(1234)
n <- 40
iterations <- 1000
lmb <- 0.2

```

### Theoretical Mean

The theoretical mean of the exponential distribution is $\lambda^{-1}$.

```{r theoretical_mean}

lmb^-1

```

### Sample plot

If we plot a sample exponential distribution with 1000 variables we see the structure of the distribution. We can also calculate the sample's mean.

```{r sample_plot}

hist(rexp(1000,lmb))
mean(rexp(1000,lmb))

```

The mean is close to the theoretical mean.

### Simulation

When we run the simulation, we see the following:

```{r simulation}

mns = NULL
for (i in 1 : iterations) mns = c(mns, mean(rexp(n, lmb)))
hist(mns)

mean(mns)

```

The distribution of the means is normally distributed and the mean of the means is very close to the theoretical mean.
